---
title: "Exercise 7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Improve the prediction function in `mnist/mnist_svd.R`

The prediction function below runs almost 6 minutes with OpenBLAS on 8 cores. It's outer loop takes a single image to classify across all 10 models. Remove the outer for loop and use a matrix of images to predict simultaneously with each of the 10 models. What is the optimal number of cores for this faster version?
```{r eval=FALSE}
predict_svdmod = function(test, models) {
  np = nrow(test)
  pred = rep(NA, np)
  mnames = names(models)
  mloss = setNames(vector("double", length(models)), mnames)
  
  for(i in 1:np) {
    y = as.numeric(test[i, ])
    for(m in mnames) {
      vt = models[[m]]$vt
      yhat = t(vt) %*% vt %*% y # memory for compute: precompute t(vt)%*%vt
      mloss[m] = sum((y - yhat)^2)/length(y)
    }
    pred[i] = mnames[which.min(mloss)]
  }
  pred
}
```
Note that the `mnist_read.R` function was updated for read-only access so it can be accessed by more than one user at a time.